\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{tcolorbox}
\usepackage{array}
\usepackage{hyperref}

\geometry{a4paper, margin=0.9in}
\hypersetup{colorlinks=true, linkcolor=blue!70!black, urlcolor=blue}

\title{\textbf{EECS 498/598-016 Computer Graphics --- Sample Final Exam}\\
\large{Winter 2026 - Answer Key \& Detailed Solutions}}
\author{}
\date{}

\begin{document}
\maketitle

\noindent
\begin{tabular}{@{}l@{\hspace{2em}}l}
\textbf{Exam Format:} & 70 minutes, closed-note, 6 questions \\
\textbf{Style:} & Conceptual understanding, no programming, no proofs, no complex equations \\
\end{tabular}

\vspace{1em}

% ============================================================
\section*{Question 1: Transformations \& Rasterization}

\textbf{1.a.} Why is the order of transformations important when composing multiple matrices?

\vspace{2cm}

\textbf{1.b.} Write the $3\times 3$ homogeneous transformation matrix for a 2D rotation by $\theta$ counter-clockwise about the origin.

\vspace{2cm}

\textbf{1.c.} Explain how Barycentric Coordinates are used in rasterization to interpolate vertex attributes across a triangle. Write the key formula.

\vspace{3cm}

\textbf{1.d.} In the Blinn-Phong shading model, what is the ``halfway vector'' $\vec{H}$ and why is it used instead of the pure reflection vector $\vec{R}$?

\vspace{3.5cm}

% ============================================================
\section*{Question 2: Ray Tracing}

\textbf{2.a.} What are the three components of the rendering equation, and what does the integral term represent physically?

\vspace{2cm}

\textbf{2.b.} Explain how recursive ray tracing simulates global illumination effects.

\vspace{2.5cm}

\textbf{2.c.} Describe the BVH tree traversal algorithm for a single ray. How does it skip unnecessary triangle intersection tests?

\vspace{2.5cm}

\textbf{2.d.} A radiometer measures $50\,\mathrm{W/m^2\!\cdot\!sr}$ of light coming from a specific direction toward a surface patch. Which radiometric quantity is this: Radiance or Irradiance? Explain the difference in terms of directionality.

\vspace{2.5cm}

% ============================================================
\section*{Question 3: Neural Fields}

\textbf{3.a.} List three advantages of neural implicit representations compared to voxels or point clouds.

\vspace{2cm}

\textbf{3.b.} Explain why the original form of neural fields is slow for training and inference. Suggest how we can make this faster.

\vspace{2.5cm}

\textbf{3.c.} For NeRF volume rendering, we don't store the opacity value directly but instead model the density $\sigma$. Explain why this choice is reasonable and beneficial.

\vspace{2.5cm}

\textbf{3.d.} For NeRF volume rendering, why do we need importance sampling for points along the ray? Explain conceptually how it can be implemented in practice.

\vspace{2.5cm}

% ============================================================
\section*{Question 4: Geometry}

\textbf{4.a.} Compare a triangular mesh and an implicit surface (SDF): which representation is better for checking if a point is inside or outside an object, and why?

\vspace{2cm}

\textbf{4.b.} Describe the property of Bézier curves regarding the convex hull of their control points.

\vspace{2cm}

\textbf{4.c.} Explain the purpose of the \textit{Marching Cubes} algorithm. What triggers triangle generation in a single grid cell?

\vspace{2.5cm}

% ============================================================
\section*{Question 5: Reconstruction}

\textbf{5.a.} Write the camera projection equation mapping a 3D world point $\mathbf{X}$ to a 2D image pixel $\mathbf{x}$.

\vspace{2cm}

\textbf{5.b.} What is the epipolar constraint equation? How does it simplify the correspondence search between two images?

\vspace{2.5cm}

\textbf{5.c.} Explain the epipolar constraint equation $\mathbf{x}'^T F \mathbf{x} = 0$. Why can it NOT be written as $\mathbf{x}^T F \mathbf{x}' = 0$?

\vspace{2cm}

\textbf{5.d.} Name two specific scenarios where Multi-View Stereo (MVS) algorithms struggle or fail.

\vspace{2cm}

% ============================================================
\section*{Question 6: Generative Models}

\textbf{6.a.} What is the main advantage of GANs over VAEs in terms of generated image quality, and what loss function difference causes this?

\vspace{2cm}

\textbf{6.b.} Explain what the KL divergence term does in the VAE loss function.

\vspace{2cm}

\textbf{6.c.} What is the purpose of Classifier-Free Guidance in diffusion models, and what is the formula?

\vspace{2.5cm}

\textbf{6.d.} How do Flow Models (Flow Matching) differ from Diffusion Models in terms of generation process and efficiency?

\vspace{2cm}

% ============================================================
\newpage
\section*{Answer Key \& Detailed Solutions}

\vspace{0.5em}

% Q1 A-D
\textbf{Q1.a.} Matrix multiplication is \textbf{non-commutative} ($AB \neq BA$). Example: translating then rotating an object rotates it around the origin at its new position, while rotating then translating moves it in a rotated direction—these produce different final positions.

\vspace{0.75em}
\textbf{Q1.b.}
$$\begin{bmatrix} x' \\ y' \\ 1 \end{bmatrix} = \begin{bmatrix} \cos\theta & -\sin\theta & 0 \\ \sin\theta & \cos\theta & 0 \\ 0 & 0 & 1 \end{bmatrix} \begin{bmatrix} x \\ y \\ 1 \end{bmatrix}$$

\vspace{0.75em}
\textbf{Q1.c.} Any point inside a triangle can be expressed as a weighted sum of its three vertices: $P = \alpha A + \beta B + \gamma C$ where $\alpha+\beta+\gamma=1$. For any vertex attribute $f$, its value at $P$ is interpolated as $f_P = \alpha f_A + \beta f_B + \gamma f_C$. The weights $\alpha,\beta,\gamma$ are computed from the relative areas of the sub-triangles.

\vspace{0.75em}
\textbf{Q1.d.} The halfway vector $\vec{H} = \frac{\vec{L}+\vec{V}}{\|\vec{L}+\vec{V}\|}$ is the normalized sum of the light direction and view direction. In the Blinn-Phong model, $\max(0, \vec{N}\cdot\vec{H})^n$ replaces the Phong reflection calculation $\max(0, \vec{N}\cdot\vec{R})^n$. $\vec{H}$ is \textbf{cheaper to compute} because it doesn't require reflecting $\vec{L}$ about the normal, and it produces similar visual results.

\vspace{1em}

\textbf{Q2.a.} The rendering equation components are:
(1) \textbf{Emitted radiance} $L_e$ – light directly produced by the surface;
(2) \textbf{Integral} $\int_\Omega$ – sums up light arriving from \textbf{all possible incoming directions} $\omega_i$ over the hemisphere;
(3) \textbf{BRDF} $f_r$ – describes how much incoming light reflects toward the outgoing direction;
(4) \textbf{Cosine term} $\mathbf{n}\cdot\omega_i$ – weights incoming light by the angle (Lambert's Law).
The integral represents the total reflected light accumulated from every direction in the scene, modeling the infinite light bounces that create global illumination.

\vspace{0.75em}
\textbf{Q2.b.} Recursive ray tracing traces secondary rays at each hit:
\begin{itemize}[leftmargin=2em, nosep]
\item \textbf{Reflection rays} bounce off shiny surfaces, producing mirror effects and interreflections between objects.
\item \textbf{Refraction rays} bend through transparent materials, creating caustics and distortion effects.
\item \textbf{Shadow rays} check visibility toward each light source for soft shadows.
\end{itemize}
By recursively bouncing these rays and accumulating their weighted contributions (capped at a max depth), the algorithm captures effects like color bleeding and soft shadows that simple local lighting cannot produce.

\vspace{0.75em}
\textbf{Q2.c.} BVH groups object triangles into a tree of bounding boxes:
\begin{enumerate}[leftmargin=2em, nosep]
\item Check if the ray intersects the root bounding box.
\item \textbf{If NO}: skip the entire subtree—none of its triangles can be hit.
\item \textbf{If YES}: recursively check child nodes.
\item At leaf nodes, test individual triangles.
\end{enumerate}
The key insight is that \textbf{ray-box intersection is much cheaper than ray-triangle intersection}. By wrapping many triangles in one box, a single cheap test can eliminate hundreds of expensive tests. This reduces complexity from $O(N)$ to roughly $O(\log N)$ for uniformly distributed geometry.

\vspace{0.75em}
\textbf{Q2.d.} This is \textbf{Radiance}. The unit $\mathrm{W/m^2\!\cdot\!sr}$ (watts per unit area per unit solid angle) is the hallmark of radiance.
\begin{itemize}[leftmargin=2em, nosep]
\item \textbf{Radiance} describes light \textbf{along a single direction}—it is the brightness of a specific beam or ray, constant along its path in a vacuum.
\item \textbf{Irradiance} describes total power arriving at a surface from \textbf{all directions} in the hemisphere above it ($\mathrm{W/m^2}$—no $sr$).
\end{itemize}
Since the measurement specifies "from a specific direction," it has directionality, making it Radiance.

\vspace{1em}

\textbf{Q3.a.} Three benefits of neural implicit representations:
\begin{enumerate}[leftmargin=2em, nosep]
\item \textbf{Resolution-free / Continuous}: Query any spatial coordinate at arbitrary precision—no pixelation, no blockiness (unlike voxels which are discrete and fixed-resolution).
\item \textbf{Memory efficient}: Compresses millions/billions of data points into megabytes of network weights (unlike point clouds which store every point individually).
\item \textbf{Differentiable}: The representation is a neural network with gradients, enabling end-to-end optimization from multi-view images and seamless integration with deep learning methods.
\end{enumerate}

\vspace{0.75em}
\textbf{Q3.b.} \textbf{Why slow}: NeRF samples dozens of points per ray per pixel. Each point requires a full MLP forward pass. An image with $1920\times 1080$ pixels and 64 samples per ray results in $\sim\!132$ \textbf{million} MLP evaluations per frame.
\\
\textbf{Faster methods}: (1) \textbf{Instant-NGP} uses a multi-resolution hash grid to store spatial features, requiring only a tiny MLP to decode them. (2) \textbf{Occupancy grids} mark empty regions of space so rays skip them entirely. (3) \textbf{3D Gaussian Splatting} replaces MLPs with splatting—orders of magnitude faster inference.

\vspace{0.75em}
\textbf{Q3.c.}
\begin{itemize}[leftmargin=2em, nosep]
\item \textbf{Density $\sigma$ is a continuous material property} independent of the sampling interval $\delta_i$. Opacity $\alpha$ depends on both $\sigma$ \textit{and} the step size: $\alpha_i = 1 - \exp(-\sigma_i \delta_i)$.
\item \textbf{Benefit}: NeRF's coarse network uses widely-spaced samples and the fine network uses densely-spaced samples. By modeling $\sigma$, both networks predict the same underlying physical property, and the formula automatically adapts $\alpha$ to each sampling rate. If the network predicted $\alpha$ directly, the values would only be valid for that specific step size.
\end{itemize}

\vspace{0.75em}
\textbf{Q3.d.} Most of the ray passes through empty space—uniform sampling wastes computation there and undersamples object surfaces, producing blurry boundaries.
\\
\textbf{Implementation (Coarse-to-Fine)}: A coarse network first estimates density along the ray with uniform sampling. This produces a probability distribution of where surface material exists. The fine network then uses \textbf{importance sampling}—drawing more sample points in regions of high predicted density (near surfaces) and fewer in empty regions, using the coarse network's PDF for stratified sampling.

\vspace{1em}

\textbf{Q4.a.} \textbf{Implicit surface (SDF)}. To check: simply evaluate $f(P)$. If $f(P)<0$, point $P$ is inside; if $f(P)>0$, $P$ is outside; if $f(P)=0$, $P$ is on the surface. This is a single function call. A mesh requires a ray-cast test from $P$ outward and counting intersections, or complex winding number calculations—both are significantly more expensive.

\vspace{0.75em}
\textbf{Q4.b.} A Bézier curve \textbf{always lies entirely within the convex hull} of its control points. The curve can never extend outside the region bounded by the control polygon.

\vspace{0.75em}
\textbf{Q4.c.} \textbf{Purpose}: Convert an implicit surface (defined by $f(x,y,z)=0$) into a triangular mesh for GPU rendering.
\\
\textbf{Trigger}: For each grid cell, evaluate $f$ at all 8 corners. \textbf{If the signs are mixed} (some corners positive, some negative), the surface passes through the cell. The algorithm interpolates zero-crossing points on the cell edges, then looks up a pre-computed table (15 cases after symmetry) to generate the corresponding triangles.

\vspace{1em}

\textbf{Q5.a.}
$$\mathbf{x} = K\,[R \mid t]\,\mathbf{X}$$
where $\mathbf{X}$ is a homogeneous 3D world point and $\mathbf{x}$ is a homogeneous 2D image pixel.

\vspace{0.75em}
\textbf{Q5.b.} $\mathbf{x}'^T F \mathbf{x} = 0$. For any point $\mathbf{x}$ in Image~1, this constrains its matching point $\mathbf{x}'$ in Image~2 to lie on a specific line (the \textbf{epipolar line} $\mathbf{l}' = F\mathbf{x}$). Instead of searching the entire 2D image for correspondences, we only search along this 1D line.

\vspace{0.75em}
\textbf{Q5.c.} Because the Fundamental Matrix maps Image~1 $\to$ Image~2 ($\mathbf{l}' = F\mathbf{x}$). The mathematically equivalent backward form uses the transpose:
$$\mathbf{x}^T F^T \mathbf{x}' = 0$$
Since $F$ is generally \textbf{not symmetric} ($F \neq F^T$), the expression $\mathbf{x}^T F \mathbf{x}'$ is NOT equal to zero.

\vspace{0.75em}
\textbf{Q5.d.} Two failure cases:
(1) \textbf{Textureless / flat-color regions} (e.g. white walls) — no distinctive features to match between images.
(2) \textbf{Specular / reflective surfaces} (e.g. mirrors, glass) — appearance changes with viewpoint, so the same 3D point looks different in different images, breaking the correspondence assumption.
(Also acceptable: occlusions, moving objects.)

\vspace{1em}

\textbf{Q6.a.} GANs produce \textbf{sharper, higher-quality images}. This is because the adversarial loss (Minimax game between Generator and Discriminator) forces the Generator to create realistic \textbf{high-frequency details} to fool the Discriminator. VAEs use pixel-wise MSE (L2) loss which averages over possible reconstructions, resulting in \textbf{blurry outputs}.

\vspace{0.75em}
\textbf{Q6.b.} The KL divergence term \textbf{regularizes the latent distribution}. It forces the encoder's output distribution $q(z\mid x)$ to stay close to a standard normal $\mathcal{N}(0, 1)$. This ensures: (1) the latent space is \textbf{continuous and smooth} (no gaps where sampling produces nonsense), (2) we can sample from $\mathcal{N}(0,1)$ and get valid generations, and (3) similar inputs map to nearby latent codes, enabling meaningful interpolation.

\vspace{0.75em}
\textbf{Q6.c.} \textbf{Classifier-Free Guidance} balances how strongly the output follows a text/label condition vs. being unconditioned—without needing a separate classifier model.
$$\hat{\epsilon} = \epsilon_u + w\!\cdot\!(\epsilon_c - \epsilon_u)$$
where $\epsilon_c$ = conditional noise prediction, $\epsilon_u$ = unconditional noise prediction, and $w$ is the guidance scale ($w=1$ is standard, $w=7.5$ is typical for strong conditioning).

\vspace{0.75em}
\textbf{Q6.d.}
\begin{itemize}[leftmargin=2em, nosep]
\item \textbf{Diffusion} uses a \textbf{stochastic differential equation (SDE)} with many small noisy steps. Typical models require hundreds to thousands of reverse steps, making sampling slow.
\item \textbf{Flow Matching} learns a deterministic \textbf{ODE velocity field}: $\frac{d\mathbf{x}}{dt} = v_\theta(\mathbf{x}, t)$. It defines a direct, clean trajectory from noise to data, typically requiring far fewer integration steps—making generation \textbf{significantly more efficient} than Diffusion.
\end{itemize}

\end{document}
